### 2012-12-02 CMPT771/survey_report/paper.tex

%File: formatting-instruction.tex
\documentclass[letterpaper]{article}
\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\frenchspacing
%\pdfinfo{
%/Title (Formatting Instructions for Authors Using LaTeX)
%/Subject (AAAI Publications)
%/Author (AAAI Press)}
\setcounter{secnumdepth}{1}  
 \begin{document}
% The file aaai.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\title{Finding communities in graphs}
\author{Vivek Mahajan\\
vmahajan@sfu.ca\\
}
\maketitle
\begin{abstract}
\begin{quote}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions. 
\end{quote}
\end{abstract}

\section{Introduction}

\section{Related Works}

\section{Proposed Idea}

\cite{test123}
\bibliographystyle{aaai} \bibliography{testbib}
\end{document}